---
title: "autocorrelation Analysis"
author: "Xieqing Yu Haoqun Cao"
date: "`r Sys.Date()`"
output: rmdformats::readthedown
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(knitr)
# library(rmdformats)
library(kableExtra)
library(dplyr)
library(tseries)
library(lubridate)
library(tibble)
library(ggplot2)
library(hms)
## Global options
options(max.print="60")
opts_knit$set(width=75)
```

```{r}
theme_ours_slanted <- function (base_size = 11, base_family = "") {
    theme(
      plot.title = element_text(color = "black", size = 16, hjust = 0.5, face = "bold"),
      axis.title = element_text(color = "black", size = 12, hjust = 0.5),
      panel.grid.major.x  = element_blank(),
      panel.grid.major.y  = element_line(color = "white"),
      panel.grid.minor  = element_blank(),
      panel.background = element_rect(fill = "#E1F0FC"),
      panel.border = element_rect(color = "steelblue4", fill = NA),
      axis.line = element_line(color = "steelblue4"),
      axis.ticks = element_line(color = "steelblue4"),
      axis.text = element_text(color = "black"),
      axis.text.x = element_text(angle = 90, hjust = 1)
      )
}
```


## plot the hypnogram to find sleep cycle
```{r}
dta <- read.csv("~/Desktop/UCInspire program summer 2023/5-min sleep new/unitef21 hypno.csv")
```

```{r}
dta$date<-as.Date(dta$time)
dta$daytime<-format(dta$time,format="%H:%M:%S")
dta$daytime<- format(as.POSIXct(dta$time),format="%H:%M:%S")
```

```{r}
for(i in 1:(nrow(dta)-1)){
  if(dta[i,]$date != dta[i+1,]$date){
    dta[i,]$value = 5
  }
}
dta[nrow(dta),]$value=5

```


```{r}
daydta <- dta %>% 
  filter(date == '2021-04-09')
daydta %>% ggplot(aes(x = daytime, y = value))+labs(x="Daytime",y="Sleeping Stage")+geom_point()+theme_ours_slanted()+ggtitle("hypnogram of unitef21 in 2021-04-09")

```

2 light 1 deep 3 REM 4 awake

2->1
2->3
2->4

3->1
3->2
3->4

1->2
1->3
1->4

4->1
4->2
4->3
4->5



## check the transition time for the same group(2->1) in one day 
````{r}
markov_chain <- daydta$value
current_state <- markov_chain[1]
current_state_duration <- 1
result <- list()
length(result)
head <- 2
tail <- 1


for (i in 2:length(markov_chain)) {
  if (markov_chain[i] == current_state) {
    current_state_duration <- current_state_duration + 1
  } else {
    # if(T){
    if(current_state == head & markov_chain[i] == tail){
      result <- append(result, list(c(current_state, current_state_duration)))
    }
    current_state <- markov_chain[i]
    current_state_duration <- 1
  }
}

# result <- append(result, list(c(current_state, current_state_duration)))
if(length(result)==0){
  result<-list(head,0)
}
result_matrix <- matrix(unlist(result), ncol = 2, byrow = TRUE)
duration = result_matrix[,2] #vector
sum=0
for(i in 1:length(duration)){
  sum=sum+duration[i]
}
sum
```
## get the list of observed dates and convert it to a dataframe
```{r}
datelist<-list()
dta$datechar<-as.character(dta$date)
for(i in 1:(nrow(dta)-1)){
  if(dta[i,]$date != dta[i+1,]$date){
    datelist <- append(datelist,dta[i,]$datechar)
  }
}
datelist<-append(datelist,dta[nrow(dta),]$datechar)
autocorr <- data.frame(matrix(unlist(datelist), nrow=length(datelist), byrow=TRUE))
colnames(autocorr)<-"datefor21"
```

## get all the sums of transition time from 2 to 1 for different dates for the number 21 pregnant woman and form into the second column of autocorr
````{r}
transtime21of21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  markov_chain <- daydta$value
  current_state <- markov_chain[1]
  current_state_duration <- 1
  result <- list()
  head <- 2
  tail <- 1
  for (j in 2:length(markov_chain)) {
    if (markov_chain[j] == current_state) {
     current_state_duration <- current_state_duration + 1
    } else {
    # if(T){
    if(current_state == head & markov_chain[j] == tail){
      result <- append(result, list(c(current_state, current_state_duration)))
    }
    current_state <- markov_chain[j]
    current_state_duration <- 1
  }
  }
  if(length(result)==0){
  result<-list(head,0)
  }
  result_matrix <- matrix(unlist(result), ncol = 2,byrow=TRUE)
  duration = result_matrix[,2] #vector
  sum=0
  for(k in 1:length(duration)){
  sum=sum+duration[k]
  }
  transtime21of21<-append(transtime21of21,sum)
}
autocorr$transtime21for21<-matrix(unlist(transtime21of21))
autocorr<-autocorr[1:486,]
autocorr_ts<-ts(autocorr$transtime21for21)
plot(autocorr_ts)
acf(autocorr_ts)
pacf(autocorr_ts)
```
## get the list of observed dates and convert it to a dataframe
```{r}
datelist<-list()
dta$datechar<-as.character(dta$date)
for(i in 1:(nrow(dta)-1)){
  if(dta[i,]$date != dta[i+1,]$date){
    datelist <- append(datelist,dta[i,]$datechar)
  }
}
datelist<-append(datelist,dta[nrow(dta),]$datechar)
autocorr <- data.frame(matrix(unlist(datelist), nrow=length(datelist), byrow=TRUE))
colnames(autocorr)<-"datefor21"
```

## get all the sums of transition time from 2 to 3 for different dates for the number 21 pregnant woman and form into the second column of autocorr
````{r}
transtime21of21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  markov_chain <- daydta$value
  current_state <- markov_chain[1]
  current_state_duration <- 1
  result <- list()
  head <- 2
  tail <- 3
  for (j in 2:length(markov_chain)) {
    if (markov_chain[j] == current_state) {
     current_state_duration <- current_state_duration + 1
    } else {
    # if(T){
    if(current_state == head & markov_chain[j] == tail){
      result <- append(result, list(c(current_state, current_state_duration)))
    }
    current_state <- markov_chain[j]
    current_state_duration <- 1
  }
  }
  if(length(result)==0){
  result<-list(head,0)
  }
  result_matrix <- matrix(unlist(result), ncol = 2,byrow=TRUE)
  duration = result_matrix[,2] #vector
  sum=0
  for(k in 1:length(duration)){
  sum=sum+duration[k]
  }
  transtime21of21<-append(transtime21of21,sum)
}
autocorr$transtime21for21<-matrix(unlist(transtime21of21))
autocorr<-autocorr[1:486,]
autocorr_ts<-ts(autocorr$transtime21for21)
plot(autocorr_ts)
acf(autocorr_ts)
```
## get the list of observed dates and convert it to a dataframe
```{r}
datelist<-list()
dta$datechar<-as.character(dta$date)
for(i in 1:(nrow(dta)-1)){
  if(dta[i,]$date != dta[i+1,]$date){
    datelist <- append(datelist,dta[i,]$datechar)
  }
}
datelist<-append(datelist,dta[nrow(dta),]$datechar)
autocorr <- data.frame(matrix(unlist(datelist), nrow=length(datelist), byrow=TRUE))
colnames(autocorr)<-"datefor21"
```

## get all the sums of transition time from 2 to 4 for different dates for the number 21 pregnant woman and form into the second column of autocorr
````{r}
transtime21of21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  markov_chain <- daydta$value
  current_state <- markov_chain[1]
  current_state_duration <- 1
  result <- list()
  head <- 2
  tail <- 4
  for (j in 2:length(markov_chain)) {
    if (markov_chain[j] == current_state) {
     current_state_duration <- current_state_duration + 1
    } else {
    # if(T){
    if(current_state == head & markov_chain[j] == tail){
      result <- append(result, list(c(current_state, current_state_duration)))
    }
    current_state <- markov_chain[j]
    current_state_duration <- 1
  }
  }
  if(length(result)==0){
  result<-list(head,0)
  }
  result_matrix <- matrix(unlist(result), ncol = 2,byrow=TRUE)
  duration = result_matrix[,2] #vector
  sum=0
  for(k in 1:length(duration)){
  sum=sum+duration[k]
  }
  transtime21of21<-append(transtime21of21,sum)
}
autocorr$transtime21for21<-matrix(unlist(transtime21of21))
autocorr<-autocorr[1:486,]
autocorr_ts<-ts(autocorr$transtime21for21)
plot(autocorr_ts)
acf(autocorr_ts)
```
## get the list of observed dates and convert it to a dataframe
```{r}
datelist<-list()
dta$datechar<-as.character(dta$date)
for(i in 1:(nrow(dta)-1)){
  if(dta[i,]$date != dta[i+1,]$date){
    datelist <- append(datelist,dta[i,]$datechar)
  }
}
datelist<-append(datelist,dta[nrow(dta),]$datechar)
autocorr <- data.frame(matrix(unlist(datelist), nrow=length(datelist), byrow=TRUE))
colnames(autocorr)<-"datefor21"
```

## get all the sums of transition time from 3 to 1 for different dates for the number 21 pregnant woman and form into the second column of autocorr
````{r}
transtime21of21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  markov_chain <- daydta$value
  current_state <- markov_chain[1]
  current_state_duration <- 1
  result <- list()
  head <- 3
  tail <- 1
  for (j in 2:length(markov_chain)) {
    if (markov_chain[j] == current_state) {
     current_state_duration <- current_state_duration + 1
    } else {
    # if(T){
    if(current_state == head & markov_chain[j] == tail){
      result <- append(result, list(c(current_state, current_state_duration)))
    }
    current_state <- markov_chain[j]
    current_state_duration <- 1
  }
  }
  if(length(result)==0){
  result<-list(head,0)
  }
  result_matrix <- matrix(unlist(result), ncol = 2,byrow=TRUE)
  duration = result_matrix[,2] #vector
  sum=0
  for(k in 1:length(duration)){
  sum=sum+duration[k]
  }
  transtime21of21<-append(transtime21of21,sum)
}
autocorr$transtime21for21<-matrix(unlist(transtime21of21))
autocorr<-autocorr[1:486,]
autocorr_ts<-ts(autocorr$transtime21for21)
plot(autocorr_ts)
acf(autocorr_ts)
```

## get the list of observed dates and convert it to a dataframe
```{r}
datelist<-list()
dta$datechar<-as.character(dta$date)
for(i in 1:(nrow(dta)-1)){
  if(dta[i,]$date != dta[i+1,]$date){
    datelist <- append(datelist,dta[i,]$datechar)
  }
}
datelist<-append(datelist,dta[nrow(dta),]$datechar)
autocorr <- data.frame(matrix(unlist(datelist), nrow=length(datelist), byrow=TRUE))
colnames(autocorr)<-"datefor21"
```

## get all the sums of transition time from 3 to 2 for different dates for the number 21 pregnant woman and form into the second column of autocorr
````{r}
transtime21of21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  markov_chain <- daydta$value
  current_state <- markov_chain[1]
  current_state_duration <- 1
  result <- list()
  head <- 3
  tail <- 2
  for (j in 2:length(markov_chain)) {
    if (markov_chain[j] == current_state) {
     current_state_duration <- current_state_duration + 1
    } else {
    # if(T){
    if(current_state == head & markov_chain[j] == tail){
      result <- append(result, list(c(current_state, current_state_duration)))
    }
    current_state <- markov_chain[j]
    current_state_duration <- 1
  }
  }
  if(length(result)==0){
  result<-list(head,0)
  }
  result_matrix <- matrix(unlist(result), ncol = 2,byrow=TRUE)
  duration = result_matrix[,2] #vector
  sum=0
  for(k in 1:length(duration)){
  sum=sum+duration[k]
  }
  transtime21of21<-append(transtime21of21,sum)
}
autocorr$transtime21for21<-matrix(unlist(transtime21of21))
autocorr<-autocorr[1:486,]
autocorr_ts<-ts(autocorr$transtime21for21)
plot(autocorr_ts)
acf(autocorr_ts)
```

## get the list of observed dates and convert it to a dataframe
```{r}
datelist<-list()
dta$datechar<-as.character(dta$date)
for(i in 1:(nrow(dta)-1)){
  if(dta[i,]$date != dta[i+1,]$date){
    datelist <- append(datelist,dta[i,]$datechar)
  }
}
datelist<-append(datelist,dta[nrow(dta),]$datechar)
autocorr <- data.frame(matrix(unlist(datelist), nrow=length(datelist), byrow=TRUE))
colnames(autocorr)<-"datefor21"
```

## get all the sums of transition time from 3 to 4 for different dates for the number 21 pregnant woman and form into the second column of autocorr
````{r}
transtime21of21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  markov_chain <- daydta$value
  current_state <- markov_chain[1]
  current_state_duration <- 1
  result <- list()
  head <- 3
  tail <- 4
  for (j in 2:length(markov_chain)) {
    if (markov_chain[j] == current_state) {
     current_state_duration <- current_state_duration + 1
    } else {
    # if(T){
    if(current_state == head & markov_chain[j] == tail){
      result <- append(result, list(c(current_state, current_state_duration)))
    }
    current_state <- markov_chain[j]
    current_state_duration <- 1
  }
  }
  if(length(result)==0){
  result<-list(head,0)
  }
  result_matrix <- matrix(unlist(result), ncol = 2,byrow=TRUE)
  duration = result_matrix[,2] #vector
  sum=0
  for(k in 1:length(duration)){
  sum=sum+duration[k]
  }
  transtime21of21<-append(transtime21of21,sum)
}
autocorr$transtime21for21<-matrix(unlist(transtime21of21))
autocorr<-autocorr[1:486,]
autocorr_ts<-ts(autocorr$transtime21for21)
plot(autocorr_ts)
acf(autocorr_ts)
```

## get the list of observed dates and convert it to a dataframe
```{r}
datelist<-list()
dta$datechar<-as.character(dta$date)
for(i in 1:(nrow(dta)-1)){
  if(dta[i,]$date != dta[i+1,]$date){
    datelist <- append(datelist,dta[i,]$datechar)
  }
}
datelist<-append(datelist,dta[nrow(dta),]$datechar)
autocorr <- data.frame(matrix(unlist(datelist), nrow=length(datelist), byrow=TRUE))
colnames(autocorr)<-"datefor21"
```

## get all the sums of transition time from 1 to 2 for different dates for the number 21 pregnant woman and form into the second column of autocorr
````{r}
transtime21of21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  markov_chain <- daydta$value
  current_state <- markov_chain[1]
  current_state_duration <- 1
  result <- list()
  head <- 1
  tail <- 2
  for (j in 2:length(markov_chain)) {
    if (markov_chain[j] == current_state) {
     current_state_duration <- current_state_duration + 1
    } else {
    # if(T){
    if(current_state == head & markov_chain[j] == tail){
      result <- append(result, list(c(current_state, current_state_duration)))
    }
    current_state <- markov_chain[j]
    current_state_duration <- 1
  }
  }
  if(length(result)==0){
  result<-list(head,0)
  }
  result_matrix <- matrix(unlist(result), ncol = 2,byrow=TRUE)
  duration = result_matrix[,2] #vector
  sum=0
  for(k in 1:length(duration)){
  sum=sum+duration[k]
  }
  transtime21of21<-append(transtime21of21,sum)
}
autocorr$transtime21for21<-matrix(unlist(transtime21of21))
autocorr<-autocorr[1:486,]
autocorr_ts<-ts(autocorr$transtime21for21)
plot(autocorr_ts)
acf(autocorr_ts)
```

## get the list of observed dates and convert it to a dataframe
```{r}
datelist<-list()
dta$datechar<-as.character(dta$date)
for(i in 1:(nrow(dta)-1)){
  if(dta[i,]$date != dta[i+1,]$date){
    datelist <- append(datelist,dta[i,]$datechar)
  }
}
datelist<-append(datelist,dta[nrow(dta),]$datechar)
autocorr <- data.frame(matrix(unlist(datelist), nrow=length(datelist), byrow=TRUE))
colnames(autocorr)<-"datefor21"
```

## get all the sums of transition time from 1 to 3 for different dates for the number 21 pregnant woman and form into the second column of autocorr
````{r}
transtime21of21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  markov_chain <- daydta$value
  current_state <- markov_chain[1]
  current_state_duration <- 1
  result <- list()
  head <- 1
  tail <- 3
  for (j in 2:length(markov_chain)) {
    if (markov_chain[j] == current_state) {
     current_state_duration <- current_state_duration + 1
    } else {
    # if(T){
    if(current_state == head & markov_chain[j] == tail){
      result <- append(result, list(c(current_state, current_state_duration)))
    }
    current_state <- markov_chain[j]
    current_state_duration <- 1
  }
  }
  if(length(result)==0){
  result<-list(head,0)
  }
  result_matrix <- matrix(unlist(result), ncol = 2,byrow=TRUE)
  duration = result_matrix[,2] #vector
  sum=0
  for(k in 1:length(duration)){
  sum=sum+duration[k]
  }
  transtime21of21<-append(transtime21of21,sum)
}
autocorr$transtime21for21<-matrix(unlist(transtime21of21))
autocorr<-autocorr[1:486,]
autocorr_ts<-ts(autocorr$transtime21for21)
plot(autocorr_ts)
acf(autocorr_ts)
```

## get the list of observed dates and convert it to a dataframe
```{r}
datelist<-list()
dta$datechar<-as.character(dta$date)
for(i in 1:(nrow(dta)-1)){
  if(dta[i,]$date != dta[i+1,]$date){
    datelist <- append(datelist,dta[i,]$datechar)
  }
}
datelist<-append(datelist,dta[nrow(dta),]$datechar)
autocorr <- data.frame(matrix(unlist(datelist), nrow=length(datelist), byrow=TRUE))
colnames(autocorr)<-"datefor21"
```

## get all the sums of transition time from 1 to 4 for different dates for the number 21 pregnant woman and form into the second column of autocorr
````{r}
transtime21of21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  markov_chain <- daydta$value
  current_state <- markov_chain[1]
  current_state_duration <- 1
  result <- list()
  head <- 1
  tail <- 4
  for (j in 2:length(markov_chain)) {
    if (markov_chain[j] == current_state) {
     current_state_duration <- current_state_duration + 1
    } else {
    # if(T){
    if(current_state == head & markov_chain[j] == tail){
      result <- append(result, list(c(current_state, current_state_duration)))
    }
    current_state <- markov_chain[j]
    current_state_duration <- 1
  }
  }
  if(length(result)==0){
  result<-list(head,0)
  }
  result_matrix <- matrix(unlist(result), ncol = 2,byrow=TRUE)
  duration = result_matrix[,2] #vector
  sum=0
  for(k in 1:length(duration)){
  sum=sum+duration[k]
  }
  transtime21of21<-append(transtime21of21,sum)
}
autocorr$transtime21for21<-matrix(unlist(transtime21of21))
autocorr<-autocorr[1:486,]
autocorr_ts<-ts(autocorr$transtime21for21)
plot(autocorr_ts)
acf(autocorr_ts)
```


## get the list of observed dates and convert it to a dataframe
```{r}
datelist<-list()
dta$datechar<-as.character(dta$date)
for(i in 1:(nrow(dta)-1)){
  if(dta[i,]$date != dta[i+1,]$date){
    datelist <- append(datelist,dta[i,]$datechar)
  }
}
datelist<-append(datelist,dta[nrow(dta),]$datechar)
autocorr <- data.frame(matrix(unlist(datelist), nrow=length(datelist), byrow=TRUE))
colnames(autocorr)<-"datefor21"
```

## get all the sums of transition time from 4 to 1 for different dates for the number 21 pregnant woman and form into the second column of autocorr
````{r}
transtime21of21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  markov_chain <- daydta$value
  current_state <- markov_chain[1]
  current_state_duration <- 1
  result <- list()
  head <- 4
  tail <- 1
  for (j in 2:length(markov_chain)) {
    if (markov_chain[j] == current_state) {
     current_state_duration <- current_state_duration + 1
    } else {
    # if(T){
    if(current_state == head & markov_chain[j] == tail){
      result <- append(result, list(c(current_state, current_state_duration)))
    }
    current_state <- markov_chain[j]
    current_state_duration <- 1
  }
  }
  if(length(result)==0){
  result<-list(head,0)
  }
  result_matrix <- matrix(unlist(result), ncol = 2,byrow=TRUE)
  duration = result_matrix[,2] #vector
  sum=0
  for(k in 1:length(duration)){
  sum=sum+duration[k]
  }
  transtime21of21<-append(transtime21of21,sum)
}
autocorr$transtime21for21<-matrix(unlist(transtime21of21))
autocorr<-autocorr[1:486,]
autocorr_ts<-ts(autocorr$transtime21for21)
plot(autocorr_ts)
acf(autocorr_ts)
```

## get the list of observed dates and convert it to a dataframe
```{r}
datelist<-list()
dta$datechar<-as.character(dta$date)
for(i in 1:(nrow(dta)-1)){
  if(dta[i,]$date != dta[i+1,]$date){
    datelist <- append(datelist,dta[i,]$datechar)
  }
}
datelist<-append(datelist,dta[nrow(dta),]$datechar)
autocorr <- data.frame(matrix(unlist(datelist), nrow=length(datelist), byrow=TRUE))
colnames(autocorr)<-"datefor21"
```

## get all the sums of transition time from 4 to 2 for different dates for the number 21 pregnant woman and form into the second column of autocorr
````{r}
transtime21of21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  markov_chain <- daydta$value
  current_state <- markov_chain[1]
  current_state_duration <- 1
  result <- list()
  head <- 4
  tail <- 2
  for (j in 2:length(markov_chain)) {
    if (markov_chain[j] == current_state) {
     current_state_duration <- current_state_duration + 1
    } else {
    # if(T){
    if(current_state == head & markov_chain[j] == tail){
      result <- append(result, list(c(current_state, current_state_duration)))
    }
    current_state <- markov_chain[j]
    current_state_duration <- 1
  }
  }
  if(length(result)==0){
  result<-list(head,0)
  }
  result_matrix <- matrix(unlist(result), ncol = 2,byrow=TRUE)
  duration = result_matrix[,2] #vector
  sum=0
  for(k in 1:length(duration)){
  sum=sum+duration[k]
  }
  transtime21of21<-append(transtime21of21,sum)
}
autocorr$transtime21for21<-matrix(unlist(transtime21of21))
autocorr<-autocorr[1:486,]
autocorr_ts<-ts(autocorr$transtime21for21)
plot(autocorr_ts)
acf(autocorr_ts)
```

## get the list of observed dates and convert it to a dataframe
```{r}
datelist<-list()
dta$datechar<-as.character(dta$date)
for(i in 1:(nrow(dta)-1)){
  if(dta[i,]$date != dta[i+1,]$date){
    datelist <- append(datelist,dta[i,]$datechar)
  }
}
datelist<-append(datelist,dta[nrow(dta),]$datechar)
autocorr <- data.frame(matrix(unlist(datelist), nrow=length(datelist), byrow=TRUE))
colnames(autocorr)<-"datefor21"
```

## get all the sums of transition time from 4 to 3 for different dates for the number 21 pregnant woman and form into the second column of autocorr
````{r}
transtime21of21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  markov_chain <- daydta$value
  current_state <- markov_chain[1]
  current_state_duration <- 1
  result <- list()
  head <- 4
  tail <- 3
  for (j in 2:length(markov_chain)) {
    if (markov_chain[j] == current_state) {
     current_state_duration <- current_state_duration + 1
    } else {
    # if(T){
    if(current_state == head & markov_chain[j] == tail){
      result <- append(result, list(c(current_state, current_state_duration)))
    }
    current_state <- markov_chain[j]
    current_state_duration <- 1
  }
  }
  if(length(result)==0){
  result<-list(head,0)
  }
  result_matrix <- matrix(unlist(result), ncol = 2,byrow=TRUE)
  duration = result_matrix[,2] #vector
  sum=0
  for(k in 1:length(duration)){
  sum=sum+duration[k]
  }
  transtime21of21<-append(transtime21of21,sum)
}
autocorr$transtime21for21<-matrix(unlist(transtime21of21))
autocorr<-autocorr[1:486,]
autocorr_ts<-ts(autocorr$transtime21for21)
plot(autocorr_ts)
acf(autocorr_ts)
```

## get the list of observed dates and convert it to a dataframe
```{r}
datelist<-list()
dta$datechar<-as.character(dta$date)
for(i in 1:(nrow(dta)-1)){
  if(dta[i,]$date != dta[i+1,]$date){
    datelist <- append(datelist,dta[i,]$datechar)
  }
}
datelist<-append(datelist,dta[nrow(dta),]$datechar)
autocorr <- data.frame(matrix(unlist(datelist), nrow=length(datelist), byrow=TRUE))
colnames(autocorr)<-"datefor21"
```

## get all the sums of transition time from 4 to 5 for different dates for the number 21 pregnant woman and form into the second column of autocorr
````{r}
transtime21of21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  markov_chain <- daydta$value
  current_state <- markov_chain[1]
  current_state_duration <- 1
  result <- list()
  head <- 4
  tail <- 5
  for (j in 2:length(markov_chain)) {
    if (markov_chain[j] == current_state) {
     current_state_duration <- current_state_duration + 1
    } else {
    # if(T){
    if(current_state == head & markov_chain[j] == tail){
      result <- append(result, list(c(current_state, current_state_duration)))
    }
    current_state <- markov_chain[j]
    current_state_duration <- 1
  }
  }
  if(length(result)==0){
  result<-list(head,0)
  }
  result_matrix <- matrix(unlist(result), ncol = 2,byrow=TRUE)
  duration = result_matrix[,2] #vector
  sum=0
  for(k in 1:length(duration)){
  sum=sum+duration[k]
  }
  transtime21of21<-append(transtime21of21,sum)
}
autocorr$transtime21for21<-matrix(unlist(transtime21of21))
autocorr<-autocorr[1:486,]
autocorr_ts<-ts(autocorr$transtime21for21)
plot(autocorr_ts)
acf(autocorr_ts)
```

## get all the duration of sleep for the number 21 pregnant woman and form into the third column of autocorr
```{r}
sleepduration21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  sleepduration21<-append(sleepduration21,nrow(daydta))
}
autocorr$daysleepdurationfor21<-matrix(unlist(sleepduration21))
autocorr_ts<-ts(autocorr$daysleepdurationfor21)
plot(autocorr_ts)
acf(autocorr_ts)
pacf(autocorr_ts)
```

## get all the REM duration for the number 21 pregnant woman and form into the fourth column of autocorr
```{r}
REMduration21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  count=0
  for(j in 1:nrow(daydta)){
    if(daydta[j,]$value==3){
      count=count+1
    }
  }
  REMduration21<-append(REMduration21,count)
}
autocorr$dayREMduration21<-matrix(unlist(REMduration21))
autocorr_ts<-ts(autocorr$dayREMduration21)
plot(autocorr_ts)
acf(autocorr_ts)
pacf(autocorr_ts)
```

## get all the deep duration for the number 21 pregnant woman and form into the fifth column of autocorr
```{r}
deepduration21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  count=0
  for(j in 1:nrow(daydta)){
    if(daydta[j,]$value==1){
      count=count+1
    }
  }
  deepduration21<-append(deepduration21,count)
}
autocorr$daydeepduration21<-matrix(unlist(deepduration21))
autocorr_ts<-ts(autocorr$daydeepduration21)
plot(autocorr_ts)
acf(autocorr_ts)
pacf(autocorr_ts)
```

## get all the light duration for the number 21 pregnant woman and form into the sixth column of autocorr
```{r}
lightduration21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  count=0
  for(j in 1:nrow(daydta)){
    if(daydta[j,]$value==2){
      count=count+1
    }
  }
  lightduration21<-append(lightduration21,count)
}
autocorr$daylightduration21<-matrix(unlist(lightduration21))
autocorr_ts<-ts(autocorr$daylightduration21)
plot(autocorr_ts)
acf(autocorr_ts)
pacf(autocorr_ts)
```

## get all the awake duration for the number 21 pregnant woman and form into the seventh column of autocorr
```{r}
awakeduration21<-list()
for(i in 1:nrow(autocorr)){
  dateval<-autocorr[i,1]
  daydta <- dta %>% 
    filter(date == dateval)
  count=0
  for(j in 1:nrow(daydta)){
    if(daydta[j,]$value==4|daydta[j,]$value==5){
      count=count+1
    }
  }
  awakeduration21<-append(awakeduration21,count)
}
autocorr$dayawakeduration21<-matrix(unlist(awakeduration21))
autocorr_ts<-ts(autocorr$dayawakeduration21)
plot(autocorr_ts)
acf(autocorr_ts)
pacf(autocorr_ts)
```
